Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2014-12-04 fib2.tex
Added an introduction and cleaned up a couple of references.

% Packages, Macros, Layout, Environments, etc.
\input{macros}
%------------------------------------------------------------------------------------------------------------%
\begin{document}

@@

%\section{Introduction}

Anyonic excitations in two-dimensional quantum systems offer tremendous promise for long-term storage and processing of quantum information~\cite{Kitaev2003}. The topological features of quantum systems that support such excitations are insensitive to local perturbations~\cite{Bravyi2010, *Bravyi2011a, *Michalakis2013}, and the excitations themselves can in general be used to perform universal quantum computation~\cite{Freedman2002, *Freedman2002b}.

This remarkable power becomes a liability when we seek to simulate these systems on classical computers.

One of the most interesting features of two-dimensional quantum systems with anyonic excitations~\cite{Wilczek1990} is their application to quantum information science, where the topological nature of the anyons offers some intrinsic protection of quantum coherence~\cite{Kitaev2003}. These topologically ordered systems are insensitive to local perturbations~\cite{Bravyi2010, *Bravyi2011a, *Michalakis2013}, have a ground-space degeneracy that is a function of the topology of the system~\cite{Wen1990, Einarsson1995} and can realize quantum gates by braiding anyons, which are naturally robust to error due to their topological nature~\cite{Kitaev2003, Nayak2008}. Many common anyon models such as the toric code~\cite{Kitaev2003} and the color codes~\cite{Bombin2006} have an Abelian structure that simplifies their analysis greatly. However, Abelian anyon models are very restricted when compared to the general class of anyonic systems. Notably, Abelian anyons lack the capacity to perform quantum computation by braiding of anyons alone. By contrast, non-Abelian anyon systems have excitations that can be used to implement topological quantum computation~\cite{Freedman2002, Freedman2002b}, and so they are of interest for quantum information processing schemes as well as quantum error correction. It is also known that the dynamics of Abelian anyons can significantly differ from those of non-Abelian anyons in some circumstances~\cite{Brennen2010, Lehman2011, Lehman2012, Zatloukal2012}.

@@

These thresholds are obtained by large-scale two-dimensional simulations of the dynamics of a dense ``gas'' of non-abelian anyons on lattice sizes of up to $48 \times 48$ nodes. All previous results have focused on quasi-one-dimensional models or single-particle quantum walks~\cite{Brennen2010, Lehman2011, Lehman2012, Zatloukal2012}.\stf{end copy \& paste from Ising.}

@@

\section{Modelling Fibonacci anyons}\label{s:model}

	A physical anyon model is abstractly described by a unitary modular tensor category~\cite{?}. This object contains the data describing the types of anyonic particles, as well as the results of topological operations such as braiding or fusion of these anyons. The Fibonacci anyon model consists of only one non-trivial particle type, conventionally called $\tau$. Denoting the vacuum by $\mathbb{I}$, we can describe the possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\mathbb{I}+\tau$, i.e.~two $\tau$ particles may either fuse to vacuum or to a $\tau$ particle.

@@

	The fusion outcome of any particular experiment depends on the history of the particles, in particular how they have braided around one another. For this reason braiding operations are considered to act unitarily on the fusion space. Pair-creation events should be considered isometries that enlarge the fusion space, while fusion events correspond to projective measurements of combined charge of some number of particles. For details of the description of these processes in Fibonacci anyons, see Ref.~\cite{?}.

@@

\section{Noise and error-correction}\label{s:noise}

@@

	\subsection{Logical errors}

@@

	\subsection{Decoding algorithm}

@@

\section{Simulation}\label{s:sim}

@@

	\subsection{Simulation algorithm}\label{s:simalg}

@@

\section{Numerical results}

@@

\section{Discussion}

>>>>

% Packages, Macros, Layout, Environments, etc.
\input{macros}

% use this for multiline block comments: \/* commented text here */
\long\def\/*#1*/{}
%\long\def\/*#1*/{#1} % uncomment this line to load the block comments

%------------------------------------------------------------------------------------------------------------%
\begin{document}

@@

Anyonic excitations in two-dimensional quantum systems~\cite{Wilczek1990} offer tremendous promise for long-term storage and processing of quantum information~\cite{Kitaev2003}. The topological features of quantum systems that support such excitations are insensitive to local perturbations~\cite{Bravyi2010, *Bravyi2011a, *Michalakis2013}, and the excitations themselves can in general be used to perform universal quantum computation~\cite{Freedman2002, *Freedman2002b}.

Quantum error correction is vital to harnessing the computational power of these topologically ordered systems. When coupled to a heat bath at any finite temperature, thermal fluctuations will create anyons that diffuse and quickly corrupt the stored quantum information~\cite{Pastawski2010}. Thus, the passive protection provided by the mass gap and low temperature must be augmented by an \emph{active} procedure.

Almost all of the sizable research effort on active quantum error correction for topological systems has focused on the case of abelian anyons~\cite{Terhal2014}. These systems are well suited to studying quantum error correction because the dynamics of excitations in abelian anyon models can be simulated numerically very efficiently, allowing lattice simulations of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. The ability to efficiently simulate these systems is a consequence of the limitation that abelian anyons cannot be used for universal quantum computation.

Recent investigations have begun to explore quantum error correction for nonabelian anyon models~\cite{Brell2013, Wootton2013, Hutter2014}. These models are particularly interesting because braiding and fusion of anyons in these models leads in general to universal quantum computation. However, the initial studies have focused on models such as Ising anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ model~\cite{Wootton2013, Hutter2014} which, while nonabelian, are not universal, a fact that was exploited to enable efficient classical simulation of error correction in these systems. The remarkable computational power of universal anyon models becomes a liability when we seek to simulate these systems on classical computers.

Here we consider a full simulation of quantum error correction in a two-dimensional lattice system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum computation~\cite{Wang2010b}. \stf{Describe main results and why they are bad ass.}

\/* % begin comment block

% This was copied from our Ising paper so that we would have some references
One of the most interesting features of two-dimensional quantum systems with anyonic excitations~\cite{Wilczek1990} is their application to quantum information science, where the topological nature of the anyons offers some intrinsic protection of quantum coherence~\cite{Kitaev2003}. These topologically ordered systems are insensitive to local perturbations~\cite{Bravyi2010, *Bravyi2011a, *Michalakis2013}, have a ground-space degeneracy that is a function of the topology of the system~\cite{Wen1990, Einarsson1995} and can realize quantum gates by braiding anyons, which are naturally robust to error due to their topological nature~\cite{Kitaev2003, Nayak2008}. Many common anyon models such as the toric code~\cite{Kitaev2003} and the color codes~\cite{Bombin2006} have an Abelian structure that simplifies their analysis greatly. However, Abelian anyon models are very restricted when compared to the general class of anyonic systems. Notably, Abelian anyons lack the capacity to perform quantum computation by braiding of anyons alone. By contrast, non-Abelian anyon systems have excitations that can be used to implement topological quantum computation~\cite{Freedman2002, Freedman2002b}, and so they are of interest for quantum information processing schemes as well as quantum error correction. It is also known that the dynamics of Abelian anyons can significantly differ from those of non-Abelian anyons in some circumstances~\cite{Brennen2010, Lehman2011, Lehman2012, Zatloukal2012}.

@@

These thresholds are obtained by large-scale two-dimensional simulations of the dynamics of a dense ``gas'' of non-abelian anyons on lattice sizes of up to $48 \times 48$ nodes. All previous results have focused on quasi-one-dimensional models or single-particle quantum walks~\cite{Brennen2010, Lehman2011, Lehman2012, Zatloukal2012}.\stf{end copy \& paste from Ising.}
*/ % end commented block

@@

\paragraph{Modelling Fibonacci anyons.}\label{s:model}

	A physical anyon model is abstractly described by a unitary modular tensor category~\cite{Wang2010b}. This object contains the data describing the types of anyonic particles, as well as the results of topological operations such as braiding or fusion of these anyons. The Fibonacci anyon model consists of only one non-trivial particle type, conventionally called $\tau$. Denoting the vacuum by $\mathbb{I}$, we can describe the possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\mathbb{I}+\tau$, i.e.~two $\tau$ particles may either fuse to vacuum or to a $\tau$ particle.

@@

	The fusion outcome of any particular experiment depends on the history of the particles, in particular how they have braided around one another. For this reason braiding operations are considered to act unitarily on the fusion space. Pair-creation events should be considered isometries that enlarge the fusion space, while fusion events correspond to projective measurements of combined charge of some number of particles. For details of the description of these processes in Fibonacci anyons, see Ref.~\cite{Trebst2008}.

@@

\paragraph{Noise and error-correction}\label{s:noise}

@@

	\paragraph{Logical errors}

@@

	\paragraph{Decoding algorithm}

@@

\paragraph{Simulation}\label{s:sim}

@@

	\paragraph{Simulation algorithm}\label{s:simalg}

@@

\paragraph{Numerical results}

@@

\paragraph{Discussion}
